Best Practices for Alchemical Free Energy Calculations: v\versionnumber

### 2020-01-20 paper/manuscript.tex
uniform use of vectors

    \Delta G_{\mathrm{bind}} \cong \Delta F= -k_BT\ln\frac{\int_\mathrm{bound} \exp(-\beta U(\mathbf{x}))d\mathbf{x}}{\int_\mathrm{unbound} \exp(-\beta U(\mathbf{x}))d\mathbf{x}},
where $\beta = \frac{1}{k_BT}$ and $F$ is the Helmholtz free energy. 
Or more generally in a simplified form the free energy difference ($\Delta f$) between the bound and unbound states is:
\Delta f \equiv f(\mathrm{bound}) - f(\mathrm{unbound}) = -\ln\frac{Z(\mathrm{unbound})}{Z({\mathrm{bound}})},
where $Z = \int_\mathrm{bound} \exp(-\beta U(\mathbf{x}))d\mathbf{x}$ represents the configurational integral. 
Unfortuantely it is not possible to evaluate the integral required for the configurational integral exactly, but can be approximated using molecular dynamics simulations. 
So, how can we compute $\Delta G_{\mathrm{bind}}$ from a molecular dynamics simulation? Running a direct simulation of the binding events is too computationally costly, meaning we cannot easily estimate the configurational integrals of the bound and unbound state from the simulation, as typically binding events happen at a $\mu$s scale, and simulations reach around 1 $\mu$s/day. For a reliable estimate of a single free energy of binding many days of simulations would be required, making these type of calculations uninteresting at a pharamaceutical timescale. Instead, we can run alchemical free energy simulations, by making use of the fact that free energy is a state function. We can write down a thermodynamic cycle as seen in ~\ref{fig:fig_binding_thermodynamic_cycle}, for a relative free energy simulation. Based on this cycle we can compute the relative free energy of binding between two compounds $A$ and $B$, by going along the arrow of $\Delta G_{\mathrm{bound}}$, where we introduce a second compound that is similar to the first one. Now $\Delta G_{\mathrm{bound}}$ can be expressed as a ratio of partition functions of toluene (A) and benzyl alcohol (B). 
The defining characteristic of alchemical free energy calculations is the use of a series of alchemically-modified potential functions $U(x; \lambda)$ in which an alchemical parameter $\lambda$ modulates interactions in a manner that cannot occur in real chemical systems.
One or more simulations are used to collect data from a multitude of alchemical states to compute a free energy difference between a chemical state ($\lambda_0$) and another chemical or alchemical reference state ($\lambda_1$),
\Delta f &\equiv& f(\lambda_1) - f(\lambda_0) = - \ln \frac{Z(\lambda_1)}{Z(\lambda_0)} , \label{equation:dimensionless-free-energy-difference}
where the dimensionless free energy $f(\lambda) \equiv \beta F(\lambda)$ is given in terms of partition functions $Z(\lambda)$,
Z(\lambda) &=& \int_{\Gamma} d\vec{q} \, e^{-u(\vec{q}; \lambda)} ,
\label{equation:partition-function-definition}
, where $\Gamma$ signifies the phase space volume over which to integrate. 
The \emph{reduced potential} $u(\vec{q}; \lambda)$~\cite{shirts2008statisticallya} is generally given by a trace over thermodynamic parameters with their conjugate dynamical variables,
u(\vec{q};\lambda) &\equiv& \beta \left[ U(\vec{q};\lambda) + p \, V(\vec{q}) + \sum_{i=1}^N \mu_i \, N_i(\vec{q}) + \cdots \right] . \label{equation:reduced-potential}
Here, the collection of thermodynamic and alchemical parameters $\theta \equiv \{\beta, \lambda, p, \mu, \ldots\}$ defines a \emph{thermodynamic state}.
The physical transformation of interest---such as a protein-ligand association process, or transfer free energy among phases---may involve a thermodynamic cycle that requires several free energy differences $\Delta f$ to be computed in order to produce an estimate of the overall desired quantity, see for example the thermodynamic cycle for an absolute free energy calculation as seen in~\ref{fig:fig_absolute_thermodynamic_cycle}.

>>>>

\begin{equation}
    \Delta G_{\mathrm{bind}} \cong \Delta F= -k_BT\ln\frac{\int_\mathrm{bound} \exp(-\beta U(\vec{q}))d\vec{q}}{\int_\mathrm{unbound} \exp(-\beta U(\vec{q}))d\vec{q}},
\end{equation}
where $\beta = \frac{1}{k_BT}$ and $F$ is the Helmholtz free energy. 
Or more generally in a simplified form the free energy difference ($\Delta f$) between the bound and unbound states is:
\begin{equation}
\Delta f \equiv f(\mathrm{bound}) - f(\mathrm{unbound}) = -\ln\frac{Z(\mathrm{unbound})}{Z({\mathrm{bound}})},
\end{equation}
where $Z = \int_\mathrm{bound} \exp(-\beta U(\vec{q}))d\vec{q}$ represents the configurational integral. 
Unfortuantely it is not possible to evaluate the integral required for the configurational integral exactly, but can be approximated using molecular dynamics simulations. 
So, how can we compute $\Delta G_{\mathrm{bind}}$ from a molecular dynamics simulation? Running a direct simulation of the binding events is too computationally costly, meaning we cannot easily estimate the configurational integrals of the bound and unbound state from the simulation, as typically binding events happen at a $\mu$s scale, and simulations reach around 1 $\mu$s/day. For a reliable estimate of a single free energy of binding many days of simulations would be required, making these type of calculations uninteresting at a pharamaceutical timescale. Instead, we can run alchemical free energy simulations, by making use of the fact that free energy is a state function. We can write down a thermodynamic cycle as seen in ~\ref{fig:fig_binding_thermodynamic_cycle}, for a relative free energy simulation. Based on this cycle we can compute the relative free energy of binding between two compounds $A$ and $B$, by going along the arrow of $\Delta G_{\mathrm{bound}}$, where we introduce a second compound that is similar to the first one. Now $\Delta G_{\mathrm{bound}}$ can be expressed as a ratio of partition functions of toluene (A) and benzyl alcohol (B). 
The defining characteristic of alchemical free energy calculations is the use of a series of alchemically-modified potential functions $U(x; \lambda)$ in which an alchemical parameter $\lambda$ modulates interactions in a manner that cannot occur in real chemical systems.
One or more simulations are used to collect data from a multitude of alchemical states to compute a free energy difference between a chemical state ($\lambda_0$) and another chemical or alchemical reference state ($\lambda_1$),
\begin{eqnarray}
\Delta f &\equiv& f(\lambda_1) - f(\lambda_0) = - \ln \frac{Z(\lambda_1)}{Z(\lambda_0)} , \label{equation:dimensionless-free-energy-difference}
\end{eqnarray}
where the dimensionless free energy $f(\lambda) \equiv \beta F(\lambda)$ is given in terms of partition functions $Z(\lambda)$,
\begin{eqnarray}
Z(\lambda) &=& \int_{\Gamma} d\vec{q} \, e^{-u(\vec{q}; \lambda)} ,
\label{equation:partition-function-definition}
\end{eqnarray}
, where $\Gamma$ signifies the phase space volume over which to integrate. 
The \emph{reduced potential} $u(\vec{q}; \lambda)$~\cite{shirts2008statisticallya} is generally given by a trace over thermodynamic parameters with their conjugate dynamical variables,
\begin{eqnarray}
u(\vec{q};\lambda) &\equiv& \beta \left[ U(\vec{q};\lambda) + p \, V(\vec{q}) + \sum_{i=1}^N \mu_i \, N_i(\vec{q}) + \cdots \right] . \label{equation:reduced-potential}
\end{eqnarray}
Here, the collection of thermodynamic and alchemical parameters $\theta \equiv \{\beta, \lambda, p, \mu, \ldots\}$ defines a \emph{thermodynamic state}.
The physical transformation of interest---such as a protein-ligand association process, or transfer free energy among phases---may involve a thermodynamic cycle that requires several free energy differences $\Delta f$ to be computed in order to produce an estimate of the overall desired quantity, see for example the thermodynamic cycle for an absolute free energy calculation as seen in~\ref{fig:fig_absolute_thermodynamic_cycle}.
